---
title: "Time Series Labb 1"
subtitle: "Time Series Analysis, Umeå University"
author: "Artem Shiryaev\\
Group 8"
date: "`r Sys.Date()`"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Problem Forumlation

- Time Series Forecasting 
- Investment Decision based on Accuracy of Models
- Exploration of data
- Model evaluation


```{r, echo = F, include = F}
load("~/University/Education/Mathematics/Avancerade Kurser/VT24 Time Series/Assignments/Labb 1/UMU_Time_Series_VT24/Workspace_labb1.RData")
load("~/University/Education/Mathematics/Avancerade Kurser/VT24 Time Series/Assignments/Labb 1/UMU_Time_Series_VT24/Workspace_labb2.RData")

```
## Data Presentation
```{r, echo = F, include = T}
par(mfrow=c(1,2))
plot(formated.data[,1],
     ylab = "Opening Prices of BTC",
     main = "Original data"
)

plot(monthly.data,
     ylab ="Opening Prices of BTC in USD",
          main = "Smoothed Monthly data")


```
```{r, echo =F, include =F}
library(forecast)

```

## Data Presentation

```{r, echo = F, include = T}
par(mfrow=c(2,2))
plot(train.data,
     ylab = "Opening Prices of BTC",
     xlab = "Years",
          main = "Split Training data"
)

plot(test.data,
     ylab = "Opening Prices of BTC",
     xlab = "Months in 2023",
               main = "Split test data"
)


plot(log.train.data,
     ylab = "log Opening Prices of BTC",
     xlab = "Years",
     main = "Log train data"
)

plot(log.test.data ,
     ylab = "log Opening Prices of BTC",
     xlab = "Months in 2023",
     main = "Log test data"
)
```


<!-- ## Statistical Methods Applied -->


<!-- - S1 Method -->

<!-- \begin{align} -->
<!-- X_t &= m_t + s_t + Y_t, \:\: t=1,\dots n,\: \text{ where }, \mathbf{E} [Y_t ] = 0\\ -->
<!-- m_t &= \frac{0.5x_{t-3} + x_{t-2} + x_{t-1} + x_{t} + x_{t+1} + x_{t+2} + 0.5x_{t+3}}{6}\\ -->
<!-- s_t &= w_k - \frac{1}{d} \Sigma_{i}^{d}w_i \: \:\: \: ,i,k = 1,2,\dots d\\ -->
<!-- d_t &= x_t - s_t\\ -->
<!-- &\text{Then re-estimate the means using the de-seasonalized data}\\ -->
<!-- \hat{m}_t &= \frac{0.5d_{t-3} + d_{t-2} + d_{t-1} + d_{t} + d_{t+1} + d_{t+2} + 0.5d_{t+3}}{6}\\ -->
<!-- \hat{Y}_t &= x_t - \hat{m}_t - s_t -->
<!-- \end{align} -->

<!-- ## Statistical Methods Applied -->


<!-- - S2 Method -->

<!-- Method S2 consist of elimination of trend and seasonal component by differencing. -->

<!-- The \textbf{lag-d} difference operator $\nabla_d$ is defined as -->

<!-- $$ \nabla_d X_t = X_t - X-{t-d} = (1-\mathcal{B}^d)X_t$$ -->
<!-- Using special property of  -->
<!-- $$ \mathcal{B} X_t = X_{t-1} $$ -->

<!-- $$ \nabla_d X_t = m_t - m_{t-d} + Y_t - Y_{t-d} $$ -->




## Statistical Methods Results: S1
```{r, echo = F, include = T}
par(mfrow=c(1,2))
plot(S1.MA.Results,
     ylab = "Log BTC/USD",
     main = "S1 Filtered Time Series")
plot(S1.MA.Filter,
     ylab = "Log BTC/USD",
     main = "Unfiltered Time Series")

```


<!-- ## Statistical Methods Results: S1 -->

<!-- Seasonal components: -->

<!-- | Months  |    Jan    |  Feb      |   Mar   |  Apr    |  May      |  June     |   July    |  Aug  |   Sept   |   Oct     |  Nov      |    Dec    | -->
<!-- |----------|-----------|-----------|-----------|----------|-----------|-----------|-----------|---------|----------|-----------|-----------|-----------| -->
<!-- | $s_t$    |  -0.34    | -0.33     |   -0.41   | -0.18   | -0.14    | -0.17    | -0.25    | -0.31    | -0.39   | -0.43    | -0.45    | -0.51    | -->



<!-- ## Statistical Methods Results: S1 -->
<!-- ```{r, echo = F, include = T} -->

<!-- par(mfrow=c(1,2)) -->
<!-- plot(Xt.ts.data2, -->
<!--      ylab = "Deseasonablized log Price BTC/USD") -->

<!-- plot(S1.DS.Results, -->
<!--      ylab = "Filtered Deseasonablized log Price BTC/USD") -->


<!-- ``` -->


## Statistical Methods Results: S1
```{r, echo = F, include = T}
par(mfrow=c(1,2))
plot(Yhat_t,
     ylab = "Residuals of Time Series")
plot(Yhat_ACF,
     ylab = "ACF",
     main = "ACF of Residuals ")


```


## Statistical Methods Results: S2
```{r, echo = F, include = T, fig.cap= "Differenced Method for de-seasonlized and de-trending Time series"}
par(mfrow=c(1,2))

plot(S2.trend.diffed,
     ylab = "Residuals",
     main = "De-season/trended Time Series")
plot(acf.trend,
     main = "ACF on Residuals Dataset")

```

## Checking Assumptions

  - Visually checking the sample autocorrelation function
  
\begin{align*}
\text{H}_0 &= \text{The Time Series is iid Noise}\\
\text{H}_1 &= \text{The Time Series is } \mathbf{\text{NOT}} \text{ iid Noise}
\end{align*}

  - Protmanteau test
  
  - Turning point test
  
  - Difference-sign test
  
  - Mann-Kendall Rank test
  
  - Augmented Dicker-Fuller test
  
  - Checking for normality
    - Histogram
    - qq plot
    - Normality test
      - Shapiro-Wilks test
      - Shapiro–Francia test
      
      
      
## Checking Assumptions
```{r, echo = F, include = T, fig.cap= "ACF for S1 and S2 methods on time series"}
par(mfrow=c(1,2))

plot(Yhat_ACF,
     main = "S1 ACF on Residuals Dataset")

plot(acf.trend,
     main = "S2 ACF on Residuals Dataset")



```



## Checking Assumptions
```{r, echo = F, include=TRUE, fig.cap= "Q-Q plots for Normality of Residuals of S1 and S2 Method"}

# Normality Plots
par(mfrow=c(1,2))
qqnorm(S1.Residuals,
       main = "S1 Normal Q-Q Plot")
qqline(S1.Residuals)
qqnorm(S2.Residuals,
       main = "S2 Normal Q-Q Plot")
qqline(S2.Residuals)

```


## Spectral Analysis
```{r, echo = F, include = T, fig.cap= "Spectral Analysis on S1 and S2 data"}
S1.Method.ts <- as.ts(S1.DS.Results)
S2.Method.ts <- as.ts(S2.trend.diffed)
# Calculating the perriodogram
par(mfrow=c(2,2))
S1.spc <- spectrum(S1.Method.ts,plot="T", main = "S1 Method - Raw Perriodogram")
S2.spc <- spectrum(S2.Method.ts,plot="T", main = "S2 Method - Raw Perriodogram")
#Extracting the frequencies at which the spectral density is estimated
S1.spx <- S1.spc$freq
S2.spx <- S2.spc$freq
#Extracting the spectral density estimates, scaled by 1/frequency
S1.spy <- S1.spc$spec
S2.spy <- S2.spc$spec


#Plotting the Spectral density
#par(mfrow=c(1,2))
plot(S1.spy~S1.spx,xlab="frequency",ylab="spectral density",type="l",main="Spectral Density: S1 Method TS")
plot(S2.spy~S2.spx,xlab="frequency",ylab="spectral density",type="l",main="Spectral Density: S2 Method TS")
```

## Fitting ARMA models

 - Check ACF for AR(p), PACF for MA(q)
 - Evaluate using AIC
 - Fit model
 - Residual Diagnostics
 
 
## Fitting ARMA models
```{r, echo = F, include = T}
# Normality checking

par(mfrow=c(1,2))

# Histogram of rescaled residuals
Mod2_res_scaled <- Mod2_res/sqrt(Model_2$sigma2)
hist(Mod2_res_scaled, main="Histogram of Rescaled residuals")

# qq-plot
qqnorm(Mod2_res_scaled)
qqline(Mod2_res_scaled)
```

## Fitting ARMA models
```{r, echo = F, include = T, fig.cap="Residual checks"}
par(mfrow=c(1,1))


checkresiduals(Mod2_res)
```

## Forecast of 2023

```{r, echo = F, include = T, fig.cap="Forecast with SARIMA(9,0,1) Model with 95 CI"}

par(mfrow=c(1,2))

plot(forecast_mod3$pred,
     ylab = "Prediction",
     ylim = c(7,12),
     lwd = 2,
     type = "b")
lines(UP_CI3, col = "red")
lines(LOW_CI3,col = "red")



plot(log.test.data,
     ylab = "Test data")

```



## Forecast of 2023
```{r, echo = F, include = T, fig.cap="Forecast with auto.arima AR(1) Model with 95 CI"}

par(mfrow=c(1,2))

plot(forecast_mod4$pred,
     ylab = "Prediction",
     ylim = c(7,12),
     lwd = 2,
     type = "b")
lines(UP_CI4, col = "red")
lines(LOW_CI4,col = "red")



plot(log.test.data,
     ylab = "Test data")

```


## Summary

Model | ARMA(1,1) |    ARMA(9,1)    |  SARMA(9,0,1) $\times$ (1,0,1)   |AR(1)|
------|----------|-----------|-----------|-----------|
MSE | 1261| 1258 | 10.1 | 3.72 |  


## Conclusion


- Poor performance, linear time series model is insufficient or poorly specified for using on this kind of data set
- Stationary assumption and normality assumption violated 
- Intuitively the assumption of the model that $s_t = s_{t+d}$ may be too strong
- Poor data handling and processing from my side
- Challenging to fit a suitable model







## Questions?


Thank you for listening.
